---
title: "Pathway mapping Using the detailed Flavonoid pathway in Wikipathways format"
author: "Atsushi Fukushima <atsushi.fukushima@riken.jp>"
date: "2016/09/21 (WED)"
output: html_document
---

```{r setup}
library(rWikiPathways)
```
# Set up
Install [rWikipathways](https://github.com/egonw/rwikipathways) package. Please look at the first example code from an excellent blog <https://www.r-bloggers.com/using-the-wikipathways-api-in-r/>.


# Pathway mapping Using the detailed flavonoid pathway (WP3620) in Wikipathways format

This is an example document for pathway mapping using metabolome data. 
Please see the [WikiPathways API](http://www.wikipathways.org/index.php/Help:WikiPathways_Webservice/API) in R.

## Listing organisms and pathways in the Wikipathways is performed with: 
```{r list of organisms and pathways}
organisms <- listOrganisms()
head(organisms)

pathways <- listPathways()
head(pathways)
```

## Focussing on Arabidopsis's pathways

You can also get the info, for example:

```{r Arabidopsis pathways}
araPathways <- listPathways(organisms[2])
head(araPathways)
```

We focus on the pathway of the detailed flavonoid pathway ([WP3620](http://www.wikipathways.org/index.php/Pathway:WP3620)) in Arabidopsis. You can access to the GPML source file for the pathway:
```{r flavonoid pathway}
gpml <- getPathway(pathway="WP3620")
info <- getPathwayInfo(pathway="WP3620")
info
```

Most use will likely be via database identifiers for metabolites, called Xrefs (also check out the [R package for BridgeDb](https://github.com/BiGCAT-UM/bridgedb-r)). Please check currently available [system code](http://developers.pathvisio.org/wiki/DatabasesMapps#Supporteddatabasesystems). Now we will get [KNApSAcK-ids](http://kanaya.naist.jp/KNApSAcK/) (here, systemCode="Cks"):

```{r metabolite identifiers}
xrefs.Cks <- getXrefList(pathway="WP3620", systemCode="Cks")
```

When gettng colored-pathways from the database, we specify the ["graphId"](http://www.wikipathways.org/index.php/Help:WikiPathways_Webservice/API#GraphId). So, you can get the info by findPathwaysByXref() function.

```{r findPathwaysByXref() function}
pathways <- findPathwaysByXref(xrefs.Cks[1], systemCode="Cks")  ## using KNApSAcK <http://kanaya.naist.jp/KNApSAcK/>
pathways[grep("WP3620", pathways$id),]

pathways.all <- findPathwaysByXref(xrefs.Cks, systemCode="Cks")
target.graphId <- pathways.all[grep("WP3620", pathways.all$id),]$graphid
target.graphId <- as.character(target.graphId)
```

As an example of colored pathways created with the rWikipathways package, you can get the visualization of some flavonoids.:
```{r visualization}
svg = getColoredPathway(pathway="WP3620",
                        graphId=c("f1f5d", "d5954", "bb888", "d7ff8", "c91b1"),
                        color="FF0000");
writeLines(svg, "WP3622_FL1.svg")
```

We focus on plant anthocyanins like A5 (C00007657), A6 (C00007658), A7 (C00007660), A8 (C00007659), A9 (C00006838), and A10 (C00007661). Regarding these abbreviations, see the review by [Saito et al. (2013)](http://www.ncbi.nlm.nih.gov/pubmed/23473981).

A3 (C00006836).
```{r cyanidine}
fls <- c("C00007657", "C00007659", "C00006838",
         "C00007661", "C00007658", "C00007660",
         "C00006836")
fls.all <- findPathwaysByXref(fls, systemCode="Cks")
fls.graphId <- as.character( fls.all[grep("WP3620", fls.all$id),]$graphid )


fls.svg = getColoredPathway(pathway="WP3620",
                        graphId=fls.graphId,
                        color="FF0000");
writeLines(fls.svg, "WP3620_antho.svg")
```



If you use the rWikipathways package in your research, please cite the below WikiPathways paper as follows:

[Kutmon, M., Riutta, A., Nunes, N., Hanspers, K., Willighagen, E. L., Bohler, A., Mélius, J., Waagmeester, A., Sinha, S. R., Miller, R., Coort, S. L., Cirillo, E., Smeets, B., Evelo, C. T., Pico, A. R.,  WikiPathways: capturing the full diversity of pathway knowledge. Nucleic Acids Research 44(Database issue): D488–D494 (2016)](http://dx.doi.org/10.1093/nar/gkv1024)
