---
title: "Pathway mapping Using the AtMetExpress overview in Wikipathways format"
author: "Atsushi Fukushima <atsushi.fukushima@riken.jp>"
date: "2016/09/21 (WED)"
output: html_document
---

```{r setup}
library(rWikiPathways)
```
# Set up
Install [rWikipathways](https://github.com/egonw/rwikipathways) package. Please look at the first example code from an excellent blog <https://www.r-bloggers.com/using-the-wikipathways-api-in-r/>.


# Pathway mapping Using the AtMetExpress overview in Wikipathways format

This is an example document for pathway mapping using metabolome data. 
Please see the WikiPathways API in R <http://www.wikipathways.org/index.php/Help:WikiPathways_Webservice/API>.

## Listing organisms and pathways in the Wikipathways is performed with: 
```{r list of organisms and pathways}
organisms <- listOrganisms()
head(organisms)

pathways <- listPathways()
head(pathways)
```

## Focussing on Arabidopsis's pathways

You can also get the info, for example:

```{r Arabidopsis pathways}
araPathways <- listPathways(organisms[2])
head(araPathways)
```

We focus on the pathway of AtMetExpress overview ([WP3622](http://www.wikipathways.org/index.php/Pathway:WP3622)). You can access to the GPML source file for the pathway:
```{r AtMetExpress overview}
gpml <- getPathway(pathway="WP3622")
info <- getPathwayInfo(pathway="WP3622")
info
```

Most use will likely be via database identifiers for metabolites, called Xrefs (also check out the R package for BridgeDb <https://github.com/BiGCAT-UM/bridgedb-r>). Please check currently available system code <http://developers.pathvisio.org/wiki/DatabasesMapps#Supporteddatabasesystems>. Now we will get HMDB-ids:

```{r metabolite identifiers}
xrefs <- getXrefList(pathway="WP3622", systemCode="Ch")
```

When gettng colored-pathways from the database, we specify the "graphId"<http://www.wikipathways.org/index.php/Help:WikiPathways_Webservice/API#GraphId>. So, you can get the info by findPathwaysByXref() function.

```{r findPathwaysByXref() function}
pathways <- findPathwaysByXref(xrefs[1], systemCode="Ch")  ## using HMDB <http://www.hmdb.ca/>
pathways[grep("WP3622", pathways$id),]

pathways.all <- findPathwaysByXref(xrefs, systemCode="Ch")
target.graphId <- pathways.all[grep("WP3622", pathways.all$id),]$graphid
target.graphId <- as.character(target.graphId)
```

As an example of colored pathways created with the rWikipathways package, you can get the visualization of sucrose and phytol.:
```{r visualization}
svg = getColoredPathway(pathway="WP3622",
                        graphId=c("be281","dfaae"),
                        color=c("FF0000", "00FF00"));
writeLines(svg, "WP3622_pathway1.svg")
```

We focus on plant polyamines like Ornithine (HMDB00214), Putrescine (HMDB01414), and Spermidine (HMDB01257).
```{r polyamnes}
poly <- c("HMDB00214", "HMDB01414", "HMDB01257")
poly.all <- findPathwaysByXref(poly, systemCode="Ch")
poly.graphId <- as.character( poly.all[grep("WP3622", poly.all$id),]$graphid )


poly.svg = getColoredPathway(pathway="WP3622",
                        graphId=poly.graphId,
                        color="FF0000");
writeLines(poly.svg, "WP3622_poly.svg")
```
<embed src="./WP3622_poly.svg" type="image/svg+xml" />


If you use the rWikipathways package in your research, please cite the below WikiPathways paper as follows:

[Kutmon, M., Riutta, A., Nunes, N., Hanspers, K., Willighagen, E. L., Bohler, A., Mélius, J., Waagmeester, A., Sinha, S. R., Miller, R., Coort, S. L., Cirillo, E., Smeets, B., Evelo, C. T., Pico, A. R.,  WikiPathways: capturing the full diversity of pathway knowledge. Nucleic Acids Research 44(Database issue): D488–D494 (2016)](http://dx.doi.org/10.1093/nar/gkv1024)
